---
title: "Live Session Unit 04 Assignment"
author: "Lance Dacy"
date: "June 7, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 04
This assignment had 4 parts:

* Navigating the FiveThirtyEight Data
* Creating a Data Summary of the Data / Export to CSV
* Codebook creation on a Git Repository
* Swirl Lesson 15 on Graphs

###Question 1: FiveThirtyEight Data (30 points)
Navigate to [FiveThirtyEight](https://github.com/rudeboybert/fivethirtyeight) on GitHub and read README.md. Perform the commands listed below:

a. Install the fivethirtyeight package.

```{r}
library("fivethirtyeight")
fileURL <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/recent-grads.csv"
```

b. In the listing of Data sets in package ‘fivethirtyeight,’ assign the 22nd data set to an object ‘df.’

```{r}
library(repmis)
df <-repmis::source_data(fileURL)
```

c. Use a more detailed list of the data sets to write out the URL in a comment to the related news story.

```{r}
newsstory <- "https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/"
```

d. Using R commands, give the dimensions and column names of this data frame.

```{r}
dim(df)
colnames(df)
```

###Question 2: Data Summary (30 points)
Use your newly assigned data frame

a. Write an R command that gives you the column names of the data frame. Right after that, write one that counts the number of columns but not rows.

```{r}
colnames(df)
ncol(df)
```

b. Generate a count of each unique major_category in the data frame. Assign it to major_count.

```{r}
library(plyr)
library(dplyr)
major_count <- data.frame(df %>% group_by(df$Major_category) %>% tally())
major_count
```

c. To make things easier to read, enter par(las=2) before your plot to make the text perpendicular to the axis. Make a barplot of major_count. 

```{r}
par(las=2)
barplot(as.matrix(major_count$n, major_count$df.Major_category), 
        main="Major Distribution", 
        xlab = "Unique Counts", ylab = "Major Category", 
        horiz=TRUE, col = "blue")
```

d. Write the fivethirtyeight data to a csv file. Make sure that it does not have row labels.

```{r}
write.csv(df, file = "df.csv", row.names=FALSE)
list.files()
dfcsv <- read.csv("df.csv", header=TRUE)
str(dfcsv)
```

###Question4: Codebook (30 points)
Simulate a codebook by completing the following:

a. Start a new repository on GitHub for your SMU MSDS homework. On your local device, make sure there is a directory for Homework at the minimum.

* GitHub Respository: https://github.com/lancedacy/6306

b. Create a README.md file which explains the purpose of the repository, the topics included, the sources for the material you post, and contact information in case of questions. Remember, the one in the root directory should be general.

* GitHub Respository: https://github.com/lancedacy/6306

c. In one (or more) of the nested directories, post your RMarkdown script, HTML file, and data from ‘fivethirtyeight.’

* GitHub Respository: https://github.com/lancedacy/6306/tree/master/Units/05/Unit05Project

d. In your RMarkdown script, please provide the link to this GitHub so the grader can see it.

* GitHub Respository: https://github.com/lancedacy/6306

###Question5: Swirl (10 points)
Complete Module 15 in the R Programming course of Swirl

* GitHub Link: https://github.com/lancedacy/6306/blob/master/Units/05/Swirl15Output.txt
